### 2018-03-06 trunk/citations/paper.tex
Version that works, produces graph

strange authors:

capinha para celular horizonte artificial iphone