### 2018-03-10 citations/paper.tex
Also pull out title words; a little work on paper

\documentclass[twocolumn]{article}
\usepackage[top=1.1in, left=0.85in, right=0.85in]{geometry}

% \usepackage{supertabular}
\usepackage{amsthm}
% \usepackage{relsize}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{code}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{textcomp}

\pagestyle{empty}

\newcommand\comment[1]{}

\newcommand\st{$^{\mathrm{st}}$}
\newcommand\nd{$^{\mathrm{nd}}$}
\newcommand\rd{$^{\mathrm{rd}}$}
\renewcommand\th{$^{\mathrm{th}}$}
\newcommand\tm{$^{\mbox{\tiny \textsc{tm}}}$}

% nice fractions
\newcommand\sfrac[2]{{}\,$^{#1}$\!/{}\!$_{#2}$}

% \newcommand\citef[1]{\addtocounter{footnote}{1}\footnotetext{\cite{#1}}\ensuremath{^{\mbox{\footnotesize [\thefootnote]}}}}

\newcommand\dirham{\protect \includegraphics[width=1em]{dirham}}

% \usepackage{ulem}
% go back to italics for emphasis, though
% \normalem

\begin{document}

% Academic Advancement Advice: Adopt A A Authorship

\title{Want to succeed in academia? Take the name A A}
\author{A.~A. \\
  (tom7.org foundation)
  \thanks{
    Copyright \copyright\ 2018 the Regents of the Wikiplia Foundation.
    Appears in SIGBOVIK 2018 with the abecedarian advantage of the
    Association for Aomputational Aeresy; {\em IEEEEEE!} press,
    Verlag--Verlag volume no.~{\tt std::numeric\_limits<float>::lowest()}.
    1 \dirham
} }

\renewcommand\>{$>$}
\newcommand\<{$<$}

\date{1 April 2018}

\maketitle

\section*{Abstract}
proof at last
    1\ \dirham

\vspace{1em}
{\noindent \small {\bf Keywords}:
  A A A A A A
}

\section{Introduction}

Those with names falling in the dead-middle of the alphabet---like as
a random example, Murphy---have suspected since early life a
disadvantage. For example, when the teacher lines up the students for
pie, in order to be completely fair and avoid bias, orders them by the
perfectly objective criterion of alphabetical order. Occasionally, the
teacher realizes that the same students always get pie first, and
corrects this bias by using {\em reverse} alphabetical order.
Sometimes the teacher uses alphabetical or reverse-alphabetical by
first name, or sometimes last name. All variations of this idea are of
course unfair, as is clear to first- and second-graders, who are
powerless to rebel without (a) status in the K-12 hierarchy (b) the
statistical language in which to formalize their objection and (c)
access to a prestigious conference with permissive publication
standards.

This paper demonstrates that even within academic paper publishing
(which one could argue is obsessed with bias, prestigious conferences,
statistical language for formalizing objections, and status in the
K-12 hierarchy), this unfairness persists. Specifically, we find that
authors whose names come alphabetically earlier have higher citation
rates.

\section{The Data}

I obtained a FISA warrant based solely on a salacious and unverified
memo funded by SIGBOVIK's political opponents. The warrant was for the
AMiner Open Academic Graph, which consists of summary information for
154,771,162 papers in about 130 gigabytes of JSON data. Then I embarked
on one of the Classic Journeys of Computer Science.

\subsection{Classic Journey of Computer Science}

The task is apparently simple: Parse 130 gigabytes of JSON data to
extract the author names and citations from the papers.

multiple hypothesized causes:

 - When doing a

300 GB of json woes

strange authors:

capinha para celular horizonte artificial iphone

\end{document}