### 2019-01-26 trunk/nand/paper.tex
Cleanup, optimize, notes

flinf flops and NaN gates
flip FLOPs and NaN gates?

Some useful primitives:

        These are called minNum and maxNum in IEE 754-2008 (5.3.1, p19)

        "If exactly one argument is NaN, they return the other.
        If both are NaN they return NaN.

        NAND  0 1     MAX  nan inf
             +---         +-------
           0 |1 1     nan |nan inf
           1 |1 0     inf |inf inf

                    HYPOT  nan inf
                          +-------
                      nan |nan inf
                      inf |inf inf

        So if 0 = nan and 1 = inf, then MAX is OR.

        We also have AND, since only inf*inf (aka 1 AND 1)
        returns inf in that truth table.

        Is AND and OR together complete? (I don't think so?)
        Do we have NOT?

YES!

Here are some:
Min(-x, -1.0) + inf
hypot(nan, max(1/x, -inf))
inf - max(x, 1)
sqrt(copysign(inf, -x))

Now direct search for some gates:
NAND: inf - max(x + y, -inf)
AND: x + y
OR: hypot(x, y)
XOR: Abs(max(y, -inf) - max(x, -inf))
     Abs(min(y, -x) + max(y, -inf))
     (Wouldn't be surprised if there's a simpler one,
     but my search does not find it before it gets
     out of hand...)

NOR: -inf / max(y, max(x, -1))

And how about if we leave out min, max? They are certainly
reasonable since they are defined in IEEE 754, but perhaps
not the best joke since they are not as math-y as e.g. hypot

NOR: sqrt(copysign(-inf, -hypot(y, x)))
(Here, hypot is just computing OR.)

So another NOT is:
sqrt(copysign(-inf, -x))  (actually can be inf)

XOR: abs(copysign(-inf, y) - copysign(-inf, x))
OR: hypot(y, x)
AND: x + y
NAND: inf - copysign(-inf, x + y)

ok cool, but copysign kinda has the same problems as min,max.
maybe without that?

the database saturates with 36 entries if we remove
copysign and min, max. They are these:

[nan, nan, nan, nan] = NaN
[nan, nan, nan, ~inf] = Minus(Neg(y), x)
[nan, nan, nan, ~2.0] = Plus(Tanh(Minus(Neg(y), x)), -1)
[nan, nan, nan, ~1.0] = Tanh(Minus(Neg(y), x))
[nan, nan, nan, 0.0] = Pow(y, Neg(x))
[nan, nan, nan, 1.0] = Exp(Pow(y, Neg(x)))
[nan, nan, nan, 2.0] = Plus(Exp(Pow(y, Neg(x))), 1)
[nan, nan, nan, inf] = Plus(y, x)
[nan, nan, ~inf, ~inf] = Neg(y)
[nan, nan, ~2.0, ~2.0] = Plus(Tanh(Neg(y)), -1)
[nan, nan, ~1.0, ~1.0] = Tanh(Neg(y))
[nan, nan, ~0.0, ~0.0] = Div(-1, y)
[nan, nan, 1.0, 1.0] = Exp(Div(-1, y))
[nan, nan, 2.0, 2.0] = Minus(1, Tanh(Neg(y)))
[nan, nan, inf, inf] = y
[nan, ~inf, nan, ~inf] = Neg(x)
[nan, ~inf, ~inf, ~inf] = Neg(Hypot(y, x))
[nan, ~2.0, nan, ~2.0] = Plus(Tanh(Neg(x)), -1)
[nan, ~2.0, ~2.0, ~2.0] = Plus(Tanh(Neg(Hypot(y, x))), -1)
[nan, ~1.0, nan, ~1.0] = Tanh(Neg(x))
[nan, ~1.0, ~1.0, ~1.0] = Tanh(Neg(Hypot(y, x)))
[nan, ~0.0, nan, ~0.0] = Div(-1, x)
[nan, ~0.0, ~0.0, ~0.0] = Div(-1, Hypot(y, x))
[nan, 1.0, nan, 1.0] = Exp(Div(-1, x))
[nan, 1.0, 1.0, 1.0] = Exp(Div(-1, Hypot(y, x)))
[nan, 2.0, nan, 2.0] = Minus(1, Tanh(Neg(x)))
[nan, 2.0, 2.0, 2.0] = Minus(1, Tanh(Neg(Hypot(y, x))))
[nan, inf, nan, inf] = x
[nan, inf, inf, inf] = Hypot(y, x)
[~inf, ~inf, ~inf, ~inf] = NegInf
[~2.0, ~2.0, ~2.0, ~2.0] = Plus(-1, -1)
[~1.0, ~1.0, ~1.0, ~1.0] = -1
[0.0, 0.0, 0.0, 0.0] = 0
[1.0, 1.0, 1.0, 1.0] = 1
[2.0, 2.0, 2.0, 2.0] = Plus(1, 1)
[inf, inf, inf, inf] = Inf