### 2019-02-09 trunk/chess/papers/weak.tex
Add huddle and swarm players

% Dramatic failure of the "suicide king" strategy against
% CCCP.
1. d4 g5 2. Bxg5 Nc6 3. Bxe7 Kxe7 4. d5 Kd6
5. dxc6+ Kc5 6. Qd6+ Kc4 7. e4#

% miraculous win for "generous" strategy as black:
% playing against "suicide king". White forces black
% to mate him:
1. a4 b5 2. d3 Nh6 3. Kd2 Ng8 4. Kc3 Nh6
5. Kd4 c6 6. Ke5 f5 7. a5 Qb6 8. Bg5 Qe3+
9. Bxe3 Na6 10. f3 Nc5 11. Ra2 Ne4 12. f4 Nd2
13. Bf2 Nf3+ 14. exf3 Ng4+ 15. Kxf5 g5 16. Nh3 e5
17. Bh4 Ba3 18. b4 e4 19. Nf2 c5 20. c4 Ba6
21. g3 Rc8 22. Be2 Kf7 23. fxe4 d5 24. Qc1 dxe4
25. Rg1 bxc4 26. Rd1 cxd3 27. Qxc5 Rhf8 28. Nd2 Bc4
29. Kxg5 Rc7 30. Bf3 Ke8 31. Qe5+ Kf7 32. Qe6+ Kg7
33. Qb6 Rxf4 34. Nb3 Bb5 35. Qb7 d2 36. Nh3 Bc1
37. Qxa7 Be2 38. Nd4 h5 39. Ra1 Nf2 40. Raxc1 Nh1
41. Rf1 d1=Q 42. Qxc7+ Kh8 43. Kg6 Bc4 44. Bxd1 e3
45. Qf7 Bb3 46. Rf3 Nf2 47. Rc6 e2 48. Rc7 Rf5
49. Rc4 Rf4 50. Nb5 e1=Q 51. Rc2 Qc3 52. g4 Rf5
53. Ng1 Qc7 54. Rcc3 Nd3 55. Rc5 Nf2 56. Nc3 Ba2
57. Qh7+

// Good site for some computer chess programs with ELO:
// http://www.computerchess.org.uk/ccrl/404/

// citation for knights distance:
// Counting the Number of Squares Reachable in k
// Knight's Moves
// Amanda M. Miller, David L. Farnsworth
// Open Journal of Discrete Mathematics, 2013, 3, 151-154
// http://dx.doi.org/10.4236/ojdm.2013.33027