### 2020-03-09 trunk/chess/papers/longest.tex
Some work on longest paper/code. Shaved one slack move off, so feeling pretty confident in the argument that this is the longest...

\documentclass[twocolumn]{article}
\usepackage[top=0.25in, left=0.25in, right=0.25in, bottom=0.25in]{geometry}

\usepackage{chessboard}
\usepackage{setspace}
\usepackage{seqsplit}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{code}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{textcomp}

% XXX search for common typo "movies"

% lets me explicitly set a. or 1. etc. as enum label
\usepackage{enumitem}

\usepackage[pdftex,
            pdfauthor={Dr. Tom Murphy VII Ph.D.},
            pdftitle={Is this the longest chess game?},
            pdfkeywords={SIGBOVIK, chess}]{hyperref}

\pagestyle{empty}

\renewcommand\comment[1]{}

\newcommand\st{$^{\mathrm{st}}$}
\newcommand\nd{$^{\mathrm{nd}}$}
\newcommand\rd{$^{\mathrm{rd}}$}
\renewcommand\th{$^{\mathrm{th}}$}
\newcommand\tm{$^{\mbox{\tiny \textsc{tm}}}$}

\newcommand\sfrac[2]{{}\,$^{#1}$\!/{}\!$_{#2}$}

\newcommand\citef[1]{\addtocounter{footnote}{1}\footnotetext{\cite{#1}}\ensuremath{^{\mbox{\footnotesize [\thefootnote]}}}}

\setchessboard{showmover=false}

% Define black versions of pieces for inline use. Gross, but it works.
\newcommand{\Pawn}[1][1.3ex]{%
\adjustbox{Trim=4.3pt 2.6pt 4.3pt 0pt,width=#1,margin=0.2ex 0ex 0.2ex 0ex}{\BlackPawnOnWhite}%
}%
\newcommand{\Rook}[1][1.58ex]{%
\adjustbox{Trim=3.2pt 2.2pt 3.2pt 0pt,width=#1,raise=0ex,margin=0.1ex 0ex 0.1ex 0ex}{\BlackRookOnWhite}%
}%
\newcommand{\Knight}[1][1.85ex]{%
\adjustbox{Trim=2.3pt 2.35pt 2.5pt 0pt,width=#1,raise=-0.03ex,margin=0.14ex 0ex 0.14ex 0ex}{\BlackKnightOnWhite}%
}%
\newcommand{\Bishop}[1][1.79ex]{%
\adjustbox{Trim=2.3pt 2pt 2.3pt 0pt,width=#1,raise=-0.12ex,margin=0.1ex 0ex 0.1ex 0ex}{\BlackBishopOnWhite}%
}%
\newcommand{\Queen}[1][2.05ex]{%
\adjustbox{Trim=1.2pt 2.2pt 1.2pt 0pt,width=#1,raise=-0.08ex,margin=0.1ex 0ex 0.1ex 0ex}{\BlackQueenOnWhite}%
}%
\newcommand{\King}[1][1.95ex]{%
\adjustbox{Trim=2pt 2pt 2pt 0pt,width=#1,raise=-0.06ex,margin=0.13ex 0ex 0.13ex 0ex}{\BlackKingOnWhite}%
}%

\begin{document}

\title{Is this the longest Chess game?}
\author{Dr.~Tom~Murphy~VII~Ph.D.\thanks{
Copyright \copyright\ 2020 the Regents of the Wikiplia
Foundation. Appears in SIGBOVIK 2015 with the {\tt ShortMoveString} of the
Association for Computational Heresy; {\em IEEEEEE!} press,
Verlag-Verlag volume no.~0x40-2A.
% XXX should match the length of the game
\textcurrency 17697
}
}

\renewcommand\>{$>$}
\newcommand\<{$<$}

\date{1 April 2020}

\maketitle

\section{Introduction}

In my experience, most chess games end in a few moves. Players who
seek longevity typically do so by playing many games of chess, rather
than a protracted XXX

Although many people ``know how to play'' chess, almost nobody fully
understands the rules of chess, most authoritatively given by
FIDE~\cite{fiderules}. (See for example Figure~\ref{fig:castle960} for
a minor chess scandal that erupted in 2019 over an obscure corner case
in the rules.) Several of these ``deep cuts'' have to do with game-ending
conditions that were introduced to keep the game from going on forever.

Many chess moves are reversible (e.g. moving the knight forward and
back to its starting position~\cite{survival}), so informal games of
chess could last forever with the players repeating a short cycle. In
AD~1561, Ruy L\'opez added the ``fifty-move rule'' to prevent infinite
games.\footnote{This rule only applied to games started after its
  introduction, so it is possible that some pre-1561 games are still
  in progress and may never end.} This rule (detailed below) ensures
that irreversible moves are regularly played, and so the game always
makes progress towards an end state. Another rule, ``threefold
repetition'' also guarantees termination as a sort of backup plan
(either of these rules would suffice on its own).

% This is good for the insurers and reinsurers of chess tournaments
% (games are guaranteed to end, rather than simply ending with
% probability 1).
So, chess is formally a finite game. This good for computer
scientists, since it means that chess has a trivial O(1) optimal
solution. This allows us to move onto other important questions, like:
What is the longest chess game? In this paper I show how to compute
such a game, and then gratuitously present all of its
% XXX check that this is the final one
17,697 
%
half-moves. Even if you are a chess expert (``chexpert''), I bet you
will be surprised at some of the corner cases in the rules that are
involved.

Speaking of rules, let's first detail the two main rules that limit the
length of the game.

\begin{figure}
  \begin{center}
    \chessboard[
      % fillarea=a3-h1,printarea=a3-h1,
      setfen=2nr1r1k/pb1n1pqp/1p2p1p1/1N1pN3/5P2/1P4P1/P2PPQ1P/2R1K1RB,
      pgfstyle=curvemove,markmoves={e1-g1},
      pgfstyle=straightmove,markmoves={g1-f1},
      showmover=false]
  \end{center}
  \caption{ ({\it Nepomniachtchi -- So, 2019.}) White to move during a
    speed Chess960 (aka ``Fischer Random'') tournament. In this
    variant, the pieces start in different positions, but castling
    rules are such that the king and rook end up on the same squares
    that they would in normal chess. As a result, it is possible for
    the king or rook to not move during castling, or for the
    destination square for the king to already be occupied by the
    rook. Attempting to castle in the position depicted, grandmaster
    Ian Nepomniachtchi first touched the rook to move it out of the
    way. However, piece-touching rules require that when castling, the
    player first moves the king (and ``Each move must be played with
    one hand only''); but how? The rook is occupying g1! One commenter
    suggested tossing the king into the air, then sliding the rook to
    f1 while the king is airborne, and then watching the king land
    dead center on its target. The arbiter required Nepo to make a
    rook move instead, but this was later appealed, and the game
    replayed. } \label{fig:castle960}
\end{figure}

% TODO: (Players are collaborating, perhaps playing chess to the death)

\subsection{The seventy-five move rule} \label{sec:75}

The ``fifty-move rule,''~\cite{wikipedia50move} as it is usually
known, requires that an irreversible move is played at a minimum pace.
For the sake of this rule, irreversible moves are considered captures
and pawn moves,%
\footnote{ Other irreversible moves include: Castling, moving a piece
  so as to lose castling rights, and declining to capture {\it en
    passant} (such capture must be made immediately, so if the option
  is not taken, it cannot be regained). The fifty-move rule could be
  soundly expanded to include these, but, that's just like, not the
  rule, man.}
%
including promotion and {\it en passant} (which is also a capturing
move) but not the movement of a previously promoted piece.
Specifically~\cite{fiderules}:

\begin{quote}
{\bf 9.3}.\quad The game is drawn, upon a correct claim by the player having
the move, if:
  \begin{enumerate}[label=\alph*.]
    \item he writes his move on his scoresheet and declares to the
      arbiter his intention to make this move, which shall result in
      the last 50 moves having been made by each player without the
      movement of any pawn and without any capture, or
    \item the last 50 consecutive moves have been made by each player
      without the movement of any pawn and without any capture.
  \end{enumerate}
\end{quote}

Note that what is provided here is the option for a player to {\em
  claim} a draw (and the two provisions essentially allow either player
to claim the draw at the moment of the 50\th\ move). If neither player
is interested in a draw, either because they think their position is
still winning, or are just trying to create the longest ever chess game,
the game legally continues. That's why what is actually relevant for this
paper is provision 9.6, which defines a draw:

\begin{quote}
  {\bf 9.6}.\quad If one or both of the following occur(s) then the
  game is drawn:

  \ldots 9.6.2. any series of at least 75 moves have been made by each
  player without the movement of any pawn and without any capture. If
  the last move resulted in checkmate, that shall take precedence.
\end{quote}

\begin{figure}[htp]
  \begin{center}
    \chessboard[setfen=8/r6n/8/8/5k2/3K4/7N/3b3Q b - - 0 1]
  \end{center}
  \caption{Black to move and mate in 545 moves (!). The position was
    found (by Zakharov and Makhnichev~\cite{tablebase7}) while
    building an endgame tablebase of all possible 7-piece positions.
    Of course, the game ends prematurely in a draw because of the
    75-move rule. } \label{fig:matein545}
\end{figure}

Draws after 75 moves (per player, so really 150 half-moves) are
compulsory.

Interestingly (at least as interesting as anything in this dubious
affair) it is known that some otherwise winning endgame positions
require more than 50 moves to execute (Figure~\ref{fig:matein545}).
The rules of chess have at various times allowed for longer timers in
such known situations, but were later simplified to the fixed 50 (and
75) move limit.

\subsection{Fivefold repetition} \label{sec:fivefold}

The 75-move rule is rarely applied in practice, but its
counterpart, ``threefold repetition''~\cite{wikipediathreefold} is often
the cause of draws in modern chess. This rule states that if the same
position appears three times, the players can claim a draw:

\begin{quote}
\begin{itemize}
  \item[9.2.2.] Positions are considered the same if and only if the
same player has the move, pieces of the same kind and colour occupy
the same squares and the possible moves of all the pieces of both
players are the same. Thus positions are not the same if:
\begin{itemize}
\item[9.2.2.1.]	at the start of the sequence a pawn could have been captured en passant
\item[9.2.2.2.]	a king had castling rights with a rook that has not been moved, but forfeited these after moving. The castling rights are lost only after the king or rook is moved.
\end{itemize}
\end{itemize}
\end{quote}

Like the 75-move rule, this rule has an optional version (upon
three repetitions) and a mandatory one in 9.6:

\begin{quote}
  [The game is a draw if \ldots] \\
  9.6.1. the same position has appeared, as in 9.2.2, at least five times.
\end{quote}

% (I seem to remember that this rule used to be about moves, not positions,
% but someone discovered infinite sequences of moves that don't repeat, right?
% I can't find this citation. Maybe can recreat those sequences.)
%
% Fascinatingly (at least as fascinating as anything in this
% questionable undertaking), this rule

\medskip
Many implementations of chess ignore these rules or treat them
incorrectly. Implementation of the seventy-five move rule simply
requires a count of how many moves have transpired since a pawn move
or capture, but programs typically do not force a draw after 75 moves.
Repetition requires more work, since the program must keep track of
each of the states reached since the last irreversible move. There are
also some corner cases, such as ambiguity as to whether the starting
position has ``appeared'' before the first move~\cite{survival}. The
ubiquitous FEN notation for describing chess positions does not even
include any information about states previously reached.

\subsection{Dead position} \label{sec:insufficient}

The informal version of this rule (``insufficient material'') states
that if neither side has enough pieces to mate the opponent (for
example, a king and bishop can never mate a bare king) then the game
is draw. Again, the formal rule is more subtle:

\begin{quote}
5.2.2. The game is drawn when a position has arisen in which neither
player can checkmate the opponent's king with any series of legal
moves. The game is said to end in a `dead position'. This immediately
ends the game \ldots
\end{quote}

This clearly includes the well-known material-based cases like king
and knight vs.~king, but it also surprisingly includes many other
specific positions, especially those with forced captures
(Figure~\ref{fig:insufficient}).

\begin{figure}[htp]
  \begin{center}
    \chessboard[setfen=8/8/5K2/8/8/8/Q7/k7 b - - 0 1]
  \end{center}
  \caption{Black to move and draw in 0 (!). Most players and even
    chexperts believe that the only legal move is Kxa2, and
    that the game then ends in a draw with ``insufficient material.''
    However, this game is {\em already over}. Since neither black
    nor white can win via any series of legal moves, by rule 5.2.2
    the game immediately ends in a `dead position'. (Although see
    Section~\ref{sec:ambiguity} for possible ambiguity in this
    rule.)} \label{fig:insufficient}
\end{figure}

The insufficient material rule is curious in that it requires
nontrivial computation to implement. In order to know whether the
position is a draw, an implementation needs to be able to decide
whether or not a series of moves that results in checkmate exists.
Note that this is not nearly as bad as normal game tree search because
the two sides can collaborate to produce the mate (it is not a
$\exists \forall \exists \forall \ldots$ but rather $\exists \exists
\exists \exists \ldots$). Still, such ``helpmates'' can still be
quite deep (dozens of moves) and are interesting enough to be a
common source of chess puzzles. Proving the non-existence of a
helpmate can be very difficult indeed (Figure~\ref{fig:nohelp}).

\begin{figure}[htp]
  \begin{center}
    \chessboard[setfen=k1bB4/8/2p1p1p1/1pPpPpP1/1P1P1P2/8/8/K1Bb4 w - - 0 1]
  \end{center}
  \caption{Thinking of implementing the rules of chess? To be correct,
    you'll need your program to be able to deduce that no helpmates are
    possible in this position and thus the game is over. Stockfish rates
    this as +0.4 for white, even searching to depth 92.
    (Position is due to user {\tt supercat} on Chess StackExchange.)
  } \label{fig:nohelp}
\end{figure}

% XXX might be a good place to talk about the self-referential "maximal"
% and "minimal" interpretations of this rule?

\medskip
For completeness, note that there are other routes to a draw
(stalemate, draw offers) which we can ignore; it is easy to avoid
these situations when generating the longest game.

\section{Generating the longest game}

It is generally not hard to avoid repeating positions, so the main
obstacle we'll face is the 75-move rule. Let's call an irreversible
move that resets the 75-move counter a {\em critical move}; this is
a pawn move or capturing move (or both). The structure of the game
will be a series of critical moves (I will call these ``critical''
moves) with a maximal sequence of pointless reversible moves in
between them. If we execute the maximum number of critical moves and
make 149 half-moves (just shy of triggering the compulsory version of
the 75-move rule) between them, then this will be fairly easily
seen as a maximal game.

In fact, for most positions, it is easy to waste 149 moves and return
to the exact same position. So, the strategy for generating the
longest game can mostly be broken into two phases: Make a game with a
maximum number of critical moves (it can also contain other moves) and
then pad that game out to maximum-length excursions in between its
critical moves.

\subsection{Maximal critical moves}

Critical moves are pawn moves and captures. There are 16 pawns, each
of which has 6 squares to move into before promoting, so this is
$16\times 6$ critical moves. There are 14 capturable pieces, plus the
16 pawns (they can be captured after promoting) so this nets an
additional $16+14$ critical moves for a total of $126$. Each critical
move can be made after a maximum of $75+74$ reversible moves, giving
$(75+74+1) \times 126 = 18900$ moves. The final move would capture the
last piece, yielding a draw due to the remaining kings being insufficient
material (``dead position''). This is the upper bound.

We will not quite be able to use the entire critical move budget. If
pawns only move forward in single steps, they will eventually get
blocked by the opposing pawn on the same file. Pawns can move
diagonally off their starting file only by capturing. We have plenty
of capturing to do anyway, so this in no problem. With four captures
per side, the pawns can be doubled, with a clear route to promotion,
like so:

\begin{center}
\chessboard[setfen=2bqkb2/1p1p1p1p/1p1p1p1p/8/8/P1P1P1P1/P1P1P1P1/2BQKB2 w - - 0 1]
\end{center}

Each of these captures is {\em both} a pawn move {\em and} a capturing
move, however. This means that we lose $4+4$ critical moves off our
total budget. $150 \times 118 = 17700$ is the new upper bound.

\paragraph{Parity.} If a critical move (such as a pawn move) is made by
white, then the first pointless excursion is made by black. The
players alternate these pointless moves until black has made 75 and
white 74. Now it is white's move, and white must make a critical move
or the game ends due to the 75-move rule. If instead we wanted black
to make the next critical move, this would happen after black has made
74 and white 74 non-critical moves. Any time this happens we lose one
move against the upper bound. So, we want to minimize the number of
times we switch which player is making the critical moves. Obviously
we must switch at least once, because both black and white must
make critical moves (upper bound $17699$).

\paragraph{Starting condition.} The first critical move should be made
by black. The starting position (with white to move) is analogous to
the situation just described, as if black has just made a critical move,
and it's white's turn. White will play 75 reversible moves, black plays
74, and the 150th move is black making the first critical move.

Note that white is quite constrained during this beginning phase,
unable to make any pawn moves. Only the white knights can escape the
back rank. When we try to insert $149$ pointless moves, we'll only be
able to move the knights and rooks, and doing this 75 times must leave
e.g. one of the knights on an opposite-colored square.\footnote{Knight
  moves always change the color of the knight's square, and same for
  rook moves constrained to the a1/b1 or g1/h1 squares.} So we have to
be a little careful about the position in which we make black's first
critical move.

Since white can only free their two knights, it will not be possible
for black to double four sets of their pawns (each requires capturing
a white piece). Executing as many critical moves as possible get us
to a position like

\begin{center}
\chessboard[setfen=r1bqkb2/8/5r2/p1pn4/5p1p/1p1p1p1p/PPPPPPPP/n3K3 w q - 0 47]
\end{center}

\ldots at which point we need to switch to making critical moves by white ({\bf Switch 1}).
White can promote their unblocked pawns,\footnote{White could promote all their pawns, but this would require.. actually maybe this is possible? Try...
  % In this game you get stuck at:
  % 1N1NN1RQ/4k3/7N/p1p2b1Q/5p1p/1p1p1p1p/1P6/3K4 b - - 0 95

  )
}
making progress to a position
like

\begin{center}
  \chessboard[setfen=2krQ3/8/7N/p1p2b2/PRPQ1p1p/PpPp1p1p/8/5K2 b - - 2 86]
\end{center}

\ldots and then we switch back to black again ({\bf Switch 2}). Black
promotes the remaining pawns, giving something like

\begin{center}
  \chessboard[setfen=2krbK2/8/7r/8/P1P5/PnP5/1n1n1n2/1n1n1n2 w - - 0 118]
\end{center}

\ldots with nothing left to do. We need to promote the white pawns
before capturing them. So we switch back to white ({\bf Switch 3}) and
promote them all. White can capture all of black's remaining pieces
during this phase of critical moves, resulting in a board like

\begin{center}
  \chessboard[setfen=N2Q1K2/N7/N7/8/8/7k/8/8 b - - 0 149]
\end{center}

\comment{
  position after above: N2Q1K2/N7/N7/8/8/7k/8/8 b - - 0 149

   proof game:
  1. Nf3 h6 2. Ng1 h5 3. Nf3 f5 4. Nc3 d5
5. Ne4 b5 6. Ne5 Rh6 7. Ng4 Rc6 8. Nh6 gxh6
9. Nf6+ exf6 10. Rb1 Na6 11. Ra1 Nb4 12. Rb1 Ne7
13. Ra1 Ng6 14. Rb1 Nf4 15. Ra1 Nh3 16. Rg1 Nxg1
17. Rb1 Nh3 18. Ra1 Nf4 19. Rb1 d4 20. Ra1 Nbd5
21. Rb1 Ne3 22. Ra1 Nxf1 23. Rb1 Ne3 24. Ra1 Nxd1
25. Rb1 Ne3 26. Ra1 Ned5 27. Rb1 Nd3+ 28. Kf1 Nxc1
29. Ra1 h4 30. Rb1 h3 31. Ra1 h5 32. Rb1 h4
33. Ra1 f4 34. Rb1 f5 35. Ra1 f3 36. Rb1 f4
37. Ra1 d3 38. Rb1 b4 39. Ra1 Nb3 40. Rb1 a6
41. Ra1 a5 42. Rb1 Rf6 43. Ra1 c6 44. Rb1 c5
45. Ra1 Nxa1 46. Ke1 b3 47. g3 Re6 48. g4 Rf6
49. g5 Re6 50. g6 Rf6 51. g7 Re6 52. g8=Q Rf6
53. Qg5 Re6 54. Qe5 Bg7 55. Qd4 Qg5 56. Kd1 Qg3
57. hxg3 Bh6 58. g4 Bg7 59. g5 Bh6 60. g6 Bg5
61. g7 Bh6 62. g8=N Bd7 63. e3 O-O-O 64. e4 Rd6
65. e5 Rc6 66. e6 Rb6 67. e7 Rc6 68. e8=Q Rg6
69. c3 Rg3 70. fxg3 Bc6 71. Nxh6 Bd7 72. g4 Bc6
73. g5 Bd7 74. g6 Bc6 75. g7 Bd7 76. g8=R Nf6
77. a3 Nd5 78. a4 Nf6 79. c4 Nd5 80. Rg6 Nc3+
81. Ke1 Nc2+ 82. Kf1 Na3 83. dxc3 Bf5 84. Rb6 Bd7
85. bxa3 Bf5 86. Rb4 axb4 87. Qf8 cxd4 88. Qe8 b2
89. Qf8 b1=N 90. Qe8 d2 91. Qf8 d1=N 92. Qe8 b3
93. Qf8 b2 94. Qe8 Nd2+ 95. Kg1 b1=N 96. Qf8 Nb2
97. Qe8 Nb3 98. Qf8 d3 99. Qe8 d2 100. Qf8 d1=N
101. Kf1 f2 102. Ke2 f1=N 103. Kf3 h2 104. Qe8 Bd7
105. Ke4 f3 106. Qf7 Nfd2+ 107. Kf4 f2 108. Kg5 f1=N
109. Kg6 h1=N 110. Kf6 h3 111. Kg6 h2 112. Kf6 Nhf2
113. Kg6 h1=R 114. Kg7 Be6 115. Kh7 Bxf7 116. Kg7 Be8
117. Kf8 Rxh6 118. a5 Rh5 119. a6 Rh4 120. a7 Rh3
121. a8=N Rh2 122. a4 Rh3 123. a5 Rh4 124. a6 Rh5
125. a7 Rh4 126. Nc7 Rh3 127. a8=N Kd7 128. Na6 Ke6
129. c5 Kf6 130. c6 Kg6 131. c7 Kh5 132. c8=N Rh2
133. c4 Rh3 134. c5 Rh2 135. c6 Rh1 136. c7 Rh2
137. Na7 Rh1 138. c8=Q Nc5 139. Qxc5+ Kh4 140. Qxf2+ Kh5
141. Qxf1 Kh6 142. Qxd1 Kh7 143. Qxd2 Kg6 144. Qxb2 Kg5
145. Qxb1 Kh4 146. Qxh1+ Kg5 147. Qc6 Kh4 148. Qxe8 Kh3
149. Qxd8
  }

But !

// The way we'll find this game is to manually generate a slow game,
// that is, one where we make make 16*6 pawn moves (but 8 are
captures) // and capture all the pieces except for a single queen, and
then // capture it, too. Then all that remains is to artificially slow
// down this game even more, by inserting 75+74 half-moves that get us
back // to the position we're already in (without repeating 5 times).
The // hypothesis is that such sequences are numerous, so we can find
// them easily (perhaps by just randomly moving a piece around, or //
perhaps by lazily generating some kind of graph structure).

\subsection{Better approach}

OK, need to rewrite the above for this superior game:

1. Nf3 b6 2. Nc3 g6 3. Ne5 g5 4. Nd5 b5
5. Nc4 bxc4 6. Nf4 gxf4 7. Rg1 c6 8. Rh1 c5
9. Rg1 f6 10. Rh1 f5 11. b3 Nc6 12. b4 Nd4
13. b5 Nb3 14. cxb3 Bb7 15. a3 Rc8 16. a4 Rc6
17. a5 Rd6 18. a6 Re6 19. axb7 Rd6 20. b8=N Re6
21. b6 Nh6 22. b7 Ng4 23. Nc6 Bh6 24. b8=Q Kf8
25. b4 Ke8 26. b5 Kf8 27. Qe5 Ke8 28. b6 Kf8
29. b7 Ke8 30. b8=N Bg5 31. h3 Bh4 32. hxg4 Bg3
33. fxg3 Qc8 34. g5 Qa6 35. Nb4 Kd8 36. Qe3 Kc7
37. g6 Kb6 38. g7 Rf6 39. g8=Q Re6 40. g4 Rf6
41. g5 Re6 42. g6 Rf6 43. g7 Re6 44. Qf8 Rf6
45. g8=N Rd6 46. Nf6 Qb5 47. Nh5 Rg8 48. g3 Rgg6
49. g4 Rge6 50. g5 Rf6 51. g6 Rc6 52. g7 Qa6
53. g8=N Qa5 54. d3 Qa6 55. d4 Qa5 56. d5 Qa6
57. dxc6 Qa5 58. c7 Qa6 59. c8=R Qa4 60. Qh3 Qa6
61. e3 Qa5 62. e4 Qa6 63. e5 Qa5 64. exf6 Qa6
65. Qg7 Qa5 66. f7 Qa6 67. Re8 Qa5 68. f8=R Qa6
69. Qdg4 Qa5 70. Nc6 Qa6 71. Ne5 Qa5 72. Nf3 Qa6
73. Nd3 Qa5+ 74. Ke2 Qa6 75. Nf2 Qa5 76. Qh8 Qa3
77. Nd1 Qa2+ 78. Nd2 Qa3 79. Kf2 Qa2 80. Kg1 Qa3
81. Rxa3 f3 82. Kh2 f2 83. Bd3 f1=R 84. Qf4 Rxf4
85. Ng7 h6 86. Be4 Rxe4 87. Ne6 f4 88. Nf6 f3
89. Ng8 f2 90. Nf6 f1=B 91. Rd3 Bxd3 92. Bb2 a6
93. Bc3 a5 94. Bb2 a4 95. Bc3 a3 96. Bb2 a2
97. Bc3 a1=R 98. Ba5+ Rxa5 99. Nf4 c3 100. Ne2 c2
101. Nf4 c1=N 102. Ne2 c4 103. Ng4 Nxe2 104. Nge3 c3
105. Nf5 c2 106. Nfe3 c1=R 107. Re1 h5 108. Qf3 h4
109. Kg2 h3+ 110. Kf2 h2 111. Qfh5 h1=R 112. Ng4 Rxg4
113. Nc4+ Bxc4 114. Nb2 d6 115. Rg1 d5 116. Rg2 d4
117. Kf3 d3 118. Kf2 d2 119. Qh2 d1=R 120. Kf3 Nc3
121. Kf2 Rg8 122. Kf3 e6 123. Kf4 e5+ 124. Kf5 e4+
125. Kf6 e3 126. Rg3 e2 127. Rg2 e1=Q 128. Nd3 Rxd3
129. Rg1 Rgxg1 130. Q8h3 Rxh3 131. Ra8 Rxa8 132. Rb8+ Rxb8
133. Qxb8+ Ka5 134. Qg3 Bf7 135. Qxe1 Rg5 136. Qxc1 Ka6
137. Qxc3 Kb6 138. Qxh3 Kc5 139. Kxf7 Kc6 140. Qxh1+ Kd7
141. Kf6 Ke8 142. Qg2 Rg8 143. Ke6 Rg7 144. Qxg7 Kd8
145. Qd7\#

Black's first phase results in this:
\chessboard[setfen=rnbqkbnr/p2pp2p/8/2p2p2/2p2p2/8/PPPPPPPP/R1BQKB1R w Qkq - 0 11]

We made two captures, and each of white's pawns can be freed with
at most one capture, and two are already clear. So this totals 8 pawn
captures, the (conjectured) minimum. (Switch 1) Now white's phase:

White's first phase then ends here:

\chessboard[setfen=4RRNQ/p2pp2p/1k6/2p2p1N/2p2pQ1/R6Q/3N4/2BN1BKR b - - 0 81]

We promoted all of white's pieces, and captured all of black's. (Switch 2).
Black again. Black promotes all his pawns:

\chessboard[setfen=4RRrQ/8/1k3K2/r7/2b5/2n5/1N4RQ/2rrq2r w - - 0 128]

... and then proceeds to capture all of white's remaining pieces:

\chessboard[setfen=1r6/8/1k3K2/8/2b5/7r/4n3/2r1q1rr w - - 0 134]

Now we switch back to white making critical moves (Switch 3). White mops up
all black's pieces with the king, which is the only piece left.

... this ends up not working; when black mates the white queen is
black's 75th move, but white has only made 74. So it should really be
white that mates black! (this is kinda like the asymmetry in the starting
condition.)

So instead at this moment...

\chessboard[setfen=1r6/8/1k3K2/8/2b5/2n4r/7Q/2r1q1rr w - - 0 133]

.. we leave the white queen alive, and capture all black's pieces for real!
then mate with the white queen

\chessboard[setfen=3k4/3Q4/4K3/8/8/8/8/8 b - - 2 145]

%8/8/2k5/4K3/7q/8/8/8 w - - 1 1
% 8846. Ke6 [72 72] Kc7 [72 73] 8847. Kf7 [73 73] Kd6 [73 74] 8848. Ke8 [74 74] Q
% e7 [74 75]Slack 0: 114 time(s)

\subsection{Inserting excursions} \label{sec:excursions}

\section{A longest game}

\begin{sloppypar}
\tiny
\setstretch{0.1}
\hyphenpenalty=1
\pretolerance=2
\tolerance=2

(a long game goes here)

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

 \noindent \input{slow}
\quad 1-0

\end{sloppypar}

\nocite{chesstego}

\bibliography{chess}{}
\bibliographystyle{plain}

\end{document}