### 2020-03-10 trunk/chess/papers/longest.tex
Clean up the critical moves section

\documentclass[twocolumn]{article}
\usepackage[top=0.25in, left=0.25in, right=0.25in, bottom=0.25in]{geometry}

\usepackage{chessboard}
\usepackage{setspace}
\usepackage{seqsplit}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{code}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{textcomp}

% XXX search for common typo "movies"

% lets me explicitly set a. or 1. etc. as enum label
\usepackage{enumitem}

\usepackage[pdftex,
            pdfauthor={Dr. Tom Murphy VII Ph.D.},
            pdftitle={Is this the longest chess game?},
            pdfkeywords={SIGBOVIK, chess}]{hyperref}

\pagestyle{empty}

\renewcommand\comment[1]{}

\newcommand\st{$^{\mathrm{st}}$}
\newcommand\nd{$^{\mathrm{nd}}$}
\newcommand\rd{$^{\mathrm{rd}}$}
\renewcommand\th{$^{\mathrm{th}}$}
\newcommand\tm{$^{\mbox{\tiny \textsc{tm}}}$}

\newcommand\sfrac[2]{{}\,$^{#1}$\!/{}\!$_{#2}$}

\newcommand\citef[1]{\addtocounter{footnote}{1}\footnotetext{\cite{#1}}\ensuremath{^{\mbox{\footnotesize [\thefootnote]}}}}

\setchessboard{showmover=false}

% Define black versions of pieces for inline use. Gross, but it works.
\newcommand{\Pawn}[1][1.3ex]{%
\adjustbox{Trim=4.3pt 2.6pt 4.3pt 0pt,width=#1,margin=0.2ex 0ex 0.2ex 0ex}{\BlackPawnOnWhite}%
}%
\newcommand{\Rook}[1][1.58ex]{%
\adjustbox{Trim=3.2pt 2.2pt 3.2pt 0pt,width=#1,raise=0ex,margin=0.1ex 0ex 0.1ex 0ex}{\BlackRookOnWhite}%
}%
\newcommand{\Knight}[1][1.85ex]{%
\adjustbox{Trim=2.3pt 2.35pt 2.5pt 0pt,width=#1,raise=-0.03ex,margin=0.14ex 0ex 0.14ex 0ex}{\BlackKnightOnWhite}%
}%
\newcommand{\Bishop}[1][1.79ex]{%
\adjustbox{Trim=2.3pt 2pt 2.3pt 0pt,width=#1,raise=-0.12ex,margin=0.1ex 0ex 0.1ex 0ex}{\BlackBishopOnWhite}%
}%
\newcommand{\Queen}[1][2.05ex]{%
\adjustbox{Trim=1.2pt 2.2pt 1.2pt 0pt,width=#1,raise=-0.08ex,margin=0.1ex 0ex 0.1ex 0ex}{\BlackQueenOnWhite}%
}%
\newcommand{\King}[1][1.95ex]{%
\adjustbox{Trim=2pt 2pt 2pt 0pt,width=#1,raise=-0.06ex,margin=0.13ex 0ex 0.13ex 0ex}{\BlackKingOnWhite}%
}%

\begin{document}

\title{Is this the longest Chess game?}
\author{Dr.~Tom~Murphy~VII~Ph.D.\thanks{
Copyright \copyright\ 2020 the Regents of the Wikiplia
Foundation. Appears in SIGBOVIK 2015 with the {\tt ShortMoveString} of the
Association for Computational Heresy; {\em IEEEEEE!} press,
Verlag-Verlag volume no.~0x40-2A.
% XXX should match the length of the game
\textcurrency 17,697
}
}

\renewcommand\>{$>$}
\newcommand\<{$<$}

\date{1 April 2020}

\maketitle

\section{Introduction}

In my experience, most chess games end in a few moves. Players who
seek longevity typically do so by playing many games of chess, rather
than a protracted XXX

(shortest game is well known)

Although many people ``know how to play'' chess, almost nobody fully
understands the rules of chess, most authoritatively given by
FIDE~\cite{fiderules}. (See for example Figure~\ref{fig:castle960} for
a minor chess scandal that erupted in 2019 over an obscure corner case
in the rules.) Several of these ``deep cuts'' have to do with game-ending
conditions that were introduced to keep the game from going on forever.

Many chess moves are reversible (e.g. moving the knight forward and
back to its starting position~\cite{survival}), so informal games of
chess could last forever with the players repeating a short cycle. In
AD~1561, Ruy L\'opez added the ``fifty-move rule'' to prevent infinite
games.\footnote{This rule only applied to games started after its
  introduction, so it is possible that some pre-1561 games are still
  in progress and may never end.} This rule (detailed below) ensures
that irreversible moves are regularly played, and so the game always
makes progress towards an end state. Another rule, ``threefold
repetition'' also guarantees termination as a sort of backup plan
(either of these rules would suffice on its own).

% This is good for the insurers and reinsurers of chess tournaments
% (games are guaranteed to end, rather than simply ending with
% probability 1).
So, chess is formally a finite game. This good for computer
scientists, since it means that chess has a trivial O(1) optimal
solution. This allows us to move onto other important questions, like:
What is the longest chess game? In this paper I show how to compute
such a game, and then gratuitously present all of its
% XXX check that this is the final one
17,697 
%
half-moves. Even if you are a chess expert (``chexpert''), I bet you
will be surprised at some of the corner cases in the rules that are
involved.

Speaking of rules, let's first detail the two main rules that limit the
length of the game.

\begin{figure}
  \begin{center}
    \chessboard[
      % fillarea=a3-h1,printarea=a3-h1,
      setfen=2nr1r1k/pb1n1pqp/1p2p1p1/1N1pN3/5P2/1P4P1/P2PPQ1P/2R1K1RB,
      pgfstyle=curvemove,markmoves={e1-g1},
      pgfstyle=straightmove,markmoves={g1-f1},
      showmover=false]
  \end{center}
  \caption{ ({\it Nepomniachtchi -- So, 2019.}) White to move during a
    speed Chess960 (aka ``Fischer Random'') tournament. In this
    variant, the pieces start in different positions, but castling
    rules are such that the king and rook end up on the same squares
    that they would in normal chess. As a result, it is possible for
    the king or rook to not move during castling, or for the
    destination square for the king to already be occupied by the
    rook. Attempting to castle in the position depicted, grandmaster
    Ian Nepomniachtchi first touched the rook to move it out of the
    way. However, piece-touching rules require that when castling, the
    player first moves the king (and ``Each move must be played with
    one hand only''); but how? The rook is occupying g1! One commenter
    suggested tossing the king into the air, then sliding the rook to
    f1 while the king is airborne, and then watching the king land
    dead center on its target. The arbiter required Nepo to make a
    rook move instead, but this was later appealed, and the game
    replayed. } \label{fig:castle960}
\end{figure}

% TODO: (Players are collaborating, perhaps playing chess to the death)

\subsection{The seventy-five move rule} \label{sec:75}

The ``fifty-move rule,''~\cite{wikipedia50move} as it is usually
known, requires that an irreversible move is played at a minimum pace.
For the sake of this rule, irreversible moves are considered captures
and pawn moves,%
\footnote{ Other irreversible moves include: Castling, moving a piece
  so as to lose castling rights, and declining to capture {\it en
    passant} (such capture must be made immediately, so if the option
  is not taken, it cannot be regained). The fifty-move rule could be
  soundly expanded to include these, but, that's just like, not the
  rule, man.}
%
including promotion and {\it en passant} (which is also a capturing
move) but not the movement of a previously promoted piece.
Specifically~\cite{fiderules}:

\begin{quote}
{\bf 9.3}.\quad The game is drawn, upon a correct claim by the player having
the move, if:
  \begin{enumerate}[label=\alph*.]
    \item he writes his move on his scoresheet and declares to the
      arbiter his intention to make this move, which shall result in
      the last 50 moves having been made by each player without the
      movement of any pawn and without any capture, or
    \item the last 50 consecutive moves have been made by each player
      without the movement of any pawn and without any capture.
  \end{enumerate}
\end{quote}

Note that what is provided here is the option for a player to {\em
  claim} a draw (and the two provisions essentially allow either player
to claim the draw at the moment of the 50\th\ move). If neither player
is interested in a draw, either because they think their position is
still winning, or are just trying to create the longest ever chess game,
the game legally continues. That's why what is actually relevant for this
paper is provision 9.6, which defines a draw:

\begin{quote}
  {\bf 9.6}.\quad If one or both of the following occur(s) then the
  game is drawn:

  \ldots 9.6.2. any series of at least 75 moves have been made by each
  player without the movement of any pawn and without any capture. If
  the last move resulted in checkmate, that shall take precedence.
\end{quote}

\begin{figure}[htp]
  \begin{center}
    \chessboard[setfen=8/r6n/8/8/5k2/3K4/7N/3b3Q b - - 0 1]
  \end{center}
  \caption{Black to move and mate in 545 moves (!). The position was
    found (by Zakharov and Makhnichev~\cite{tablebase7}) while
    building an endgame tablebase of all possible 7-piece positions.
    Of course, the game ends prematurely in a draw because of the
    75-move rule. } \label{fig:matein545}
\end{figure}

Draws after 75 moves (per player, so really 150 half-moves) are
compulsory.

Interestingly (at least as interesting as anything in this dubious
affair) it is known that some otherwise winning endgame positions
require more than 50 moves to execute (Figure~\ref{fig:matein545}).
The rules of chess have at various times allowed for longer timers in
such known situations, but were later simplified to the fixed 50 (and
75) move limit.

\subsection{Fivefold repetition} \label{sec:fivefold}

The 75-move rule is rarely applied in practice, but its
counterpart, ``threefold repetition''~\cite{wikipediathreefold} is often
the cause of draws in modern chess. This rule states that if the same
position appears three times, the players can claim a draw:

\begin{quote}
\begin{itemize}
  \item[9.2.2.] Positions are considered the same if and only if the
same player has the move, pieces of the same kind and colour occupy
the same squares and the possible moves of all the pieces of both
players are the same. Thus positions are not the same if:
\begin{itemize}
\item[9.2.2.1.]	at the start of the sequence a pawn could have been captured en passant
\item[9.2.2.2.]	a king had castling rights with a rook that has not been moved, but forfeited these after moving. The castling rights are lost only after the king or rook is moved.
\end{itemize}
\end{itemize}
\end{quote}

Like the 75-move rule, this rule has an optional version (upon
three repetitions) and a mandatory one in 9.6:

\begin{quote}
  [The game is a draw if \ldots] \\
  9.6.1. the same position has appeared, as in 9.2.2, at least five times.
\end{quote}

% (I seem to remember that this rule used to be about moves, not positions,
% but someone discovered infinite sequences of moves that don't repeat, right?
% I can't find this citation. Maybe can recreat those sequences.)
%
% Fascinatingly (at least as fascinating as anything in this
% questionable undertaking), this rule

\medskip
Many implementations of chess ignore these rules or treat them
incorrectly. Implementation of the seventy-five move rule simply
requires a count of how many moves have transpired since a pawn move
or capture, but programs typically do not force a draw after 75 moves.
Repetition requires more work, since the program must keep track of
each of the states reached since the last irreversible move. There are
also some corner cases, such as ambiguity as to whether the starting
position has ``appeared'' before the first move~\cite{survival}. The
ubiquitous FEN notation for describing chess positions does not even
include any information about states previously reached.

\subsection{Dead position} \label{sec:insufficient}

The informal version of this rule (``insufficient material'') states
that if neither side has enough pieces to mate the opponent (for
example, a king and bishop can never mate a bare king) then the game
is draw. Again, the formal rule is more subtle:

\begin{quote}
5.2.2. The game is drawn when a position has arisen in which neither
player can checkmate the opponent's king with any series of legal
moves. The game is said to end in a `dead position'. This immediately
ends the game \ldots
\end{quote}

This clearly includes the well-known material-based cases like king
and knight vs.~king, but it also surprisingly includes many other
specific positions, especially those with forced captures
(Figure~\ref{fig:insufficient}).

\begin{figure}[htp]
  \begin{center}
    \chessboard[setfen=8/8/5K2/8/8/8/Q7/k7 b - - 0 1]
  \end{center}
  \caption{Black to move and draw in 0 (!). Most players and even
    chexperts believe that the only legal move is Kxa2, and
    that the game then ends in a draw with ``insufficient material.''
    However, this game is {\em already over}. Since neither black
    nor white can win via any series of legal moves, by rule 5.2.2
    the game immediately ends in a `dead position'. (Although see
    Section~\ref{sec:ambiguity} for possible ambiguity in this
    rule.)} \label{fig:insufficient}
\end{figure}

The insufficient material rule is curious in that it requires
nontrivial computation to implement. In order to know whether the
position is a draw, an implementation needs to be able to decide
whether or not a series of moves that results in checkmate exists.
Note that this is not nearly as bad as normal game tree search because
the two sides can collaborate to produce the mate (it is not a
$\exists \forall \exists \forall \ldots$ but rather $\exists \exists
\exists \exists \ldots$). Still, such ``helpmates'' can still be
quite deep (dozens of moves) and are interesting enough to be a
common source of chess puzzles. Proving the non-existence of a
helpmate can be very difficult indeed (Figure~\ref{fig:nohelp}).

\begin{figure}[htp]
  \begin{center}
    \chessboard[setfen=k1bB4/8/2p1p1p1/1pPpPpP1/1P1P1P2/8/8/K1Bb4 w - - 0 1]
  \end{center}
  \caption{Thinking of implementing the rules of chess? To be correct,
    you'll need your program to be able to deduce that no helpmates are
    possible in this position and thus the game is over. Stockfish rates
    this as +0.4 for white, even searching to depth 92.
    (Position is due to user {\tt supercat} on Chess StackExchange.)
  } \label{fig:nohelp}
\end{figure}

% XXX might be a good place to talk about the self-referential "maximal"
% and "minimal" interpretations of this rule?
% Need to have a  \label{sec:ambiguity} somewhere.

\medskip
For completeness, note that there are other routes to a draw
(stalemate, draw offers) which we can ignore; it is easy to avoid
these situations when generating the longest game.

\section{Generating the longest game}

It is generally not hard to avoid repeating positions, so the main
obstacle we'll face is the 75-move rule. Let's call an irreversible
move that resets the 75-move counter a {\em critical move}; this is
a pawn move or capturing move (or both). The structure of the game
will be a series of critical moves (I will call these ``critical''
moves) with a maximal sequence of pointless reversible moves in
between them. If we execute the maximum number of critical moves and
make 149 half-moves (just shy of triggering the compulsory version of
the 75-move rule) between them, then this will be fairly easily
seen as a maximal game.

In fact, for most positions, it is easy to waste 149 moves and return
to the exact same position. So, the strategy for generating the
longest game can mostly be broken into two phases: Make a game with a
maximum number of critical moves (it can also contain other moves) and
then pad that game out to maximum-length excursions in between its
critical moves.

\subsection{Maximal critical moves}

Critical moves are pawn moves and captures. There are 16 pawns, each
of which has 6 squares to move into before promoting, so this is
$16\times 6$ critical moves. There are 14 capturable pieces, plus the
16 pawns (they can be captured after promoting) so this nets an
additional $16+14$ critical moves for a total of $126$. Each critical
move can be made after a maximum of $75+74$ reversible moves, giving
$(75+74+1) \times 126 = 18900$ moves. The final move would capture the
last piece, yielding a draw due to the remaining kings being insufficient
material (``dead position''). This is the upper bound.

We will not quite be able to use the entire critical move budget. If
pawns only move forward in single steps, they will eventually get
blocked by the opposing pawn on the same file. Pawns can move
diagonally off their starting file only by capturing. We have plenty
of capturing to do anyway, so this in no problem. With four captures
per side, the pawns can be doubled, with a clear route to promotion,
like in Figure~\ref{fig:doubled}.

\begin{figure}[h]
\begin{center}
  \chessboard[setfen=2bqkb2/1p1p1p1p/1p1p1p1p/8/8/P1P1P1P1/P1P1P1P1/2BQKB2 w - - 0 1]
  \caption{One way to clear the promotion routes for all pawns with
    eight captures. } \label{fig:doubled}
\end{center}
\end{figure}

Each of these captures is {\em both} a pawn move {\em and} a capturing
move, however. This means that we lose $4+4$ critical moves off our
total budget. $150 \times (126 - 8) = 17700$ is the new upper bound.

\paragraph{Parity.} If a critical move (such as a pawn move) is made by
white, then the first non-critical move is made by black. The
players alternate these pointless moves until black has made $75$ and
white $74$. Now it is white's move, and white must make a critical move
or the game ends due to the 75-move rule. If instead we wanted black
to make the next critical move, this would happen after black has made
$74$ and white $74$ non-critical moves. Any time this happens we lose one
move against the upper bound. So, we want to minimize the number of
times we switch which player is making the critical moves. Obviously
we must switch at least once, because both black and white must
make critical moves. This reduces the upper bound to $17699$.

\paragraph{Starting condition.} The first critical move should be made
by black. The starting position (with white to move) is analogous to
the situation just described, as if black has just made a critical move,
and it's white's turn. White will play $75$ reversible moves, black plays
$74$, and the 150\th\ move is black making the first critical move.

Note that white is quite constrained during this beginning phase, as
pawn moves are critical moves and must be avoided. Only the white
knights can escape the back rank. When we try to insert $149$
pointless moves, we'll only be able to move the knights and rooks, and
doing this $75$ times must leave e.g. one of the knights on an
opposite-colored square.\footnote{Knight moves always change the color
  of the knight's square, and same for rook moves constrained to the
  a1/b1 or g1/h1 squares.} So we have to be a little careful about the
position in which we make black's first critical move.

Since white can only free their two knights, these are the only pieces
that can be captured by black pawns. So it will not be possible for
black to double four sets of their pawns as in
Figure~\ref{fig:doubled}. This would require white to have a phase of
critical moves to free moves to capture, then black again to finish
doubling pawns, then white again to promote its freed pawns. Each switch
costs move off the theoretical max. We can be more efficient with an
asymmetric approach.

Black's first phase of critical moves instead results in this:

\comment{
1. Nf3 b6 2. Nc3 g6 3. Ne5 g5 4. Nd5 b5
5. Nc4 bxc4 6. Nf4 gxf4 7. Rg1 c6 8. Rh1 c5
9. Rg1 f6 10. Rh1 f5 
}
\begin{center}
\chessboard[setfen=rnbqkbnr/p2pp2p/8/2p2p2/2p2p2/8/PPPPPPPP/R1BQKB1R w Qkq - 0 11]
\end{center}

The white b and g pawns have a clear route to promotion. White can
free each of the remaining 6 pawns with a single capture. Between this
and black's two pawn captures, this is the minimum of 8 pawn moves
that are also captures. Since black has plenty of freed pieces, white
can promote all of their pawns during their own phase of critical
moves, resulting in:

\comment{
11. b3 Nc6 12. b4 Nd4
13. b5 Nb3 14. cxb3 Bb7 15. a3 Rc8 16. a4 Rc6
17. a5 Rd6 18. a6 Re6 19. axb7 Rd6 20. b8=N Re6
21. b6 Nh6 22. b7 Ng4 23. Nc6 Bh6 24. b8=Q Kf8
25. b4 Ke8 26. b5 Kf8 27. Qe5 Ke8 28. b6 Kf8
29. b7 Ke8 30. b8=N Bg5 31. h3 Bh4 32. hxg4 Bg3
33. fxg3 Qc8 34. g5 Qa6 35. Nb4 Kd8 36. Qe3 Kc7
37. g6 Kb6 38. g7 Rf6 39. g8=Q Re6 40. g4 Rf6
41. g5 Re6 42. g6 Rf6 43. g7 Re6 44. Qf8 Rf6
45. g8=N Rd6 46. Nf6 Qb5 47. Nh5 Rg8 48. g3 Rgg6
49. g4 Rge6 50. g5 Rf6 51. g6 Rc6 52. g7 Qa6
53. g8=N Qa5 54. d3 Qa6 55. d4 Qa5 56. d5 Qa6
57. dxc6 Qa5 58. c7 Qa6 59. c8=R Qa4 60. Qh3 Qa6
61. e3 Qa5 62. e4 Qa6 63. e5 Qa5 64. exf6 Qa6
65. Qg7 Qa5 66. f7 Qa6 67. Re8 Qa5 68. f8=R Qa6
69. Qdg4 Qa5 70. Nc6 Qa6 71. Ne5 Qa5 72. Nf3 Qa6
73. Nd3 Qa5+ 74. Ke2 Qa6 75. Nf2 Qa5 76. Qh8 Qa3
77. Nd1 Qa2+ 78. Nd2 Qa3 79. Kf2 Qa2 80. Kg1 Qa3
81. Rxa3
}
\begin{center}
  \chessboard[setfen=4RRNQ/p2pp2p/1k6/2p2p1N/2p2pQ1/R6Q/3N4/2BN1BKR b - - 0 81]
\end{center}

Now black can promote all of its pawns and capture white's pieces.
We actually leave the white queen; this turns out to be essential:

\comment{
\ldots f3 82. Kh2 f2 83. Bd3 f1=R 84. Qf4 Rxf4
85. Ng7 h6 86. Be4 Rxe4 87. Ne6 f4 88. Nf6 f3
89. Ng8 f2 90. Nf6 f1=B 91. Rd3 Bxd3 92. Bb2 a6
93. Bc3 a5 94. Bb2 a4 95. Bc3 a3 96. Bb2 a2
97. Bc3 a1=R 98. Ba5+ Rxa5 99. Nf4 c3 100. Ne2 c2
101. Nf4 c1=N 102. Ne2 c4 103. Ng4 Nxe2 104. Nge3 c3
105. Nf5 c2 106. Nfe3 c1=R 107. Re1 h5 108. Qf3 h4
109. Kg2 h3+ 110. Kf2 h2 111. Qfh5 h1=R 112. Ng4 Rxg4
113. Nc4+ Bxc4 114. Nb2 d6 115. Rg1 d5 116. Rg2 d4
117. Kf3 d3 118. Kf2 d2 119. Qh2 d1=R 120. Kf3 Nc3
121. Kf2 Rg8 122. Kf3 e6 123. Kf4 e5+ 124. Kf5 e4+
125. Kf6 e3 126. Rg3 e2 127. Rg2 e1=Q 128. Nd3 Rxd3
129. Rg1 Rgxg1 130. Q8h3 Rxh3 131. Ra8 Rxa8 132. Rb8+ Rxb8
}

\begin{center}
  \chessboard[setfen=1r6/8/1k3K2/8/2b5/2n4r/7Q/2r1q1rr w - - 0 133]
\end{center}

Finally, white captures all of black's pieces, and mates the black king.

\comment{
133. Qxb8+ Ka5 134. Qg3 Bf7 135. Qxe1 Rg5 136. Qxc1 Ka6
137. Qxc3 Kb6 138. Qxh3 Kc5 139. Kxf7 Kc6 140. Qxh1+ Kd7
141. Kf6 Ke8 142. Qg2 Rg8 143. Ke6 Rg7 144. Qxg7 Kd8
145. Qd7\#
}

\begin{center}
  \chessboard[setfen=3k4/3Q4/4K3/8/8/8/8/8 b - - 2 145]
\end{center}

(this is three switches)

\paragraph{Ending condition.} The way this game ends is subtle for
several reasons. First, note that we left the white queen on the board
and used it for mate. It is required that white be the one mating for
parity reasons, similar to the reason black must make the first critical
move. Since white makes critical moves in the last phase, black leads
on non-critical moves; at the moment white makes the checkmating move,
black has made 75 non-critical moves, and white 74. White's 75th move
mates.

But doesn't this trigger the 75-move rule? No, this rule (9.6.2;
Section~\ref{sec:75}) has a special exception for checkmate: ``If the
last move resulted in checkmate, that shall take precedence.''
Essentially, we can treat checkmate as a type of critical move.

Why not capture the white queen too? This would be a critical move,
but once two kings remain, the game ends immediately in a draw. So
there is no length advantage here over checkmate. In fact we could do
worse: Consider the position in Figure~\ref{fig:insufficient} where
black is {\em forced} to capture the queen. This game is over prior to
the capture, so this is shorter than the checkmating sequence! Even if
the king had an escape square, it is arguable
(Section~\ref{sec:ambiguity}) that the game is over in any case: There
is no legal sequence of moves that leads to mate. Either the king
captures the queen (and then clearly no mate is possible with just the
two kings) or the king escapes, but in doing so plays the
75\th\ non-critical move, and triggers a draw by the 75-move rule.

The most foolproof way to ensure that mate is always possible is for the
game itself to end in mate. This sidesteps any ambiguity about the way
the 75-move condition should be interpreted, as well.

\subsection{Inserting excursions} \label{sec:excursions}

TODO

\section{Future work}

Some puzzles?

\section{A longest game}

\begin{sloppypar}
\tiny
\setstretch{0.1}
\hyphenpenalty=1
\pretolerance=2
\tolerance=2

(a long game goes here)

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

 \noindent \input{slow}
\quad 1-0

\end{sloppypar}

\nocite{chesstego}

\bibliography{chess}{}
\bibliographystyle{plain}

\end{document}